Supporting Information for\\ \emph{Effectiveness of UNAIDS targets and HIV vaccination across 127 countries}

### 2016-12-13 Documents/supporting_information.tex
More work on PNAS submission.

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{tikz}

\usepackage{enumitem}
\setlist{topsep=1ex}

\usepackage[font={sf},
            labelfont={bf},
            labelsep=space,
            singlelinecheck=off]{caption}
\usepackage[font={sf}]{floatrow}
\usepackage{tabularx}

\usepackage[blocks]{authblk}
\renewcommand{\Authsep}{, }
\renewcommand{\Authand}{ \& }
\renewcommand{\Authands}{, }
\renewcommand{\Affilfont}{\small}

@@

% \usepackage[style=nature]{biblatex}
% \addbibresource{supporting_information.bib}
\usepackage[numbers, round, comma, sort&compress]{natbib}
\usepackage{hypernat}

\usepackage{xr}
\externaldocument{extended_data_fig_1}
\externaldocument{extended_data_fig_2}
\externaldocument{extended_data_fig_3}

\newcommand{\comment}[1]{\textbf{[#1]}}
\newcommand{\md}{\mathrm{d}}
\newcommand{\mT}{\mathrm{T}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\mat}[1]{\mathbf{#1}}
\newcommand{\me}{\mathrm{e}}
\DeclareMathOperator{\Uniform}{Uniform}
\DeclareMathOperator{\Triangular}{Triangular}
\DeclareMathOperator{\Lognormal}{Lognormal}
\DeclareMathOperator{\BetaPERT}{Beta-PERT}

\renewcommand{\thesection}{S\arabic{section}}
\newcommand{\sectionname}{Section}
\renewcommand{\sectionautorefname}{\sectionname}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

@@

\author{Jan Medlock}
\affil{Department of Biomedical Sciences, Oregon State University, 106
  Dryden Hall, Corvallis, OR, 97331-4801, USA,
  \href{mailto:jan.medlock@oregonstate.edu}
  {\texttt{jan.medlock@oregonstate.edu}}}
\author{Abhishek Pandey}
\author{Alyssa S.~Parpia}
\author{Amber Tang}
\author{Laura A.~Skrip}
\author{Alison P.~Galvani}
\affil{Center for Infectious Disease Modeling and Analysis, Yale
  School of Public Health, 135 College Street, New Haven, USA}

@@

We developed a mathematical model of HIV transmission and progression
that stratifies HIV infection into acute, chronic undiagnosed, chronic
diagnosed, chronic treated, chronic virally suppressed, and AIDS,
along with uninfected unvaccinated and vaccinated people
(\autoref{model} \& Extended Data Fig.~\ref*{model_diag}), tailored to
each of 127 countries.  Simulations of our model project the number of
people in each HIV stratum from 2015 to 2035, from which statistics
such as prevalence and incidence were calculated.  To evaluate the
impact of the 90--90--90 and 95--95--95 UNAIDS goals, model transition
rates were adjusted dynamically to meet intervention goals for
diagnosis, treatment, viral suppression, and vaccination coverage
(\autoref{targets}).  Data and estimates from many sources, including
The World Bank and UNAIDS, were used to parametrize the model
(\autoref{data_sources}).  Country-specific transmission rates were
fitted to estimates of the historical trajectories of incidence and
prevalence, spanning from as early as 1990 for some countries
(\autoref{model_fitting}). For every scenario of intervention
combinations, we conducted 1000 model simulations, sampling values of
model parameters from empirical distributions for each simulation, and
reported the results with median and percentiles
(\autoref{uncertainty}).

The model simulation and analysis tools, written in Python, are
publicly available.\cite{medlock2016-git}

@@

Within each country, our model divides people aged $15$\;years and
older into non-overlapping HIV states.  We focused on this age group
because it accounts for the vast majority of global
PLHIV,\cite{UNICEF} 93\% in 2014; because population-level HIV
estimates and demographic data are typically only estimated for ages
$15$--$49$; and to capture the continuing burden of infections in
people ages $50$ and older.  Our model has 8 HIV states (Extended Data
Fig.~\ref*{model_diag}): susceptible to HIV infection ($S$ is the
number of people susceptible), vaccinated against HIV ($R$), acute HIV
infection ($A$), undiagnosed HIV infection ($U$), diagnosed but
untreated HIV infection ($D$), treated without viral suppression
($T$), viral suppression ($V$), and having AIDS ($W$).

@@

\begin{table}
  \begin{center}
    \begin{tabularx}{\textwidth}{lXlll}
      \hline
      & Definition & Value & References \\
      \hline
      $\mu$ & Background death rate
      & Country specific, Data Table 2
      & \cite{World_Development_Indicators2013-ee} \\
      $\kappa$ & Recruitment rate
      & Country specific, Data Table 2
      & \cite{World_Development_Indicators2013-ee, WorldBankpg} \\
      $\delta$	& Rate of progression from acute infection
      & $\Triangular$(2, 4.14, 9.6)\;y$^{-1}$
      & \cite{Hollingsworth2008-iy} \\
      $\sigma$	& Rate of developing AIDS without viral suppression
      & 0.1064\;y$^{-1}$ & \cite{Morgan2002-cq} \\
      $\theta$ & Rate of developing AIDS with viral suppression
      & Eq.~(\ref{theta}) & --- \\
      $\gamma$ & Rate of viral suppression
      & $\Uniform$(0.5, 1.5)\;y$^{-1}$
      & \cite{Currie2009-yz} \\
      $\nu$ & Death rate from AIDS & 0.5\;y$^{-1}$
      & \cite{Morgan2002-cq} \\
      $\rho$ & Vaccine efficacy & 50\%\,[30\%, 70\%] & --- \\
      $\omega$	& Reduction in life expectancy with viral suppression
      & $\Uniform$(5, 8)\;y
      & \cite{Samji2013-kf, Unaids2014-ue} \\
      $\tau_{A}$ & Transmissibility during acute phase
      & $\Triangular$(0.0039, 0.0082, 0.0150)
      & \cite{Wawer2005-us, Skarbinski2015-ni}\\
      $\tau_{U}$ & Transmissibility after acute phase
      & $\Triangular$(0.00077, 0.0014, 0.00251)
      & \cite{Hughes2012-so} \\
      $\epsilon$ & Relative transmissibility with
      viral suppression & $\BetaPERT$(0.08, 0.002, 0.57)
      & \cite{Donnell2010-xo, attia_2009, wilson_2012, jia_2013,
        rodger_2016} \\
      $n$ & Coital acts per year & $\Uniform$(96, 108)
      & \cite{Wawer2005-us, Abdool_Karim2010-cm}\\
      $\alpha$ & Diagnosis rate & See eq.~(\ref{diagnosis_rate}) & --- \\
      $\phi$ & Treatment rate & See eq.~(\ref{treatment_rate}) & --- \\
      $\psi$ & Rate of relapse to untreated & See eq.~(\ref{relapse_rate})
      & --- \\
      $\zeta$ & Vaccination rate & See eq.~(\ref{vaccination_rate}) & --- \\
      \hline
      \vspace*{-2.5em}
    \end{tabularx}
    \caption{Model parameters. $\Triangular$, $\Uniform$, and
      $\BetaPERT$ are sampling distributions (\autoref{uncertainty}).
      For vaccine efficacy, $50\%$ was the baseline, and $30\%$ and
      $70\%$ were also assessed in sensitivity analysis.}
    \label{model_param}
  \end{center}
\end{table}

People susceptible to HIV ($S$) are at risk of becoming infected
according to the force of infection ($\lambda$), which depends on the
HIV transmission rate, estimated from the country-specific incidence
and prevalence data (\autoref{model_fitting}), adjusted for the
current number of acutely infected, untreated or ineffectively
treated, and virally suppressed PLHIV.  Upon infection, people
transition into acute infection ($A$), characterized by high viral
loads for an average of 2.90 [95\% confidence interval 1.23,
6.00]\;months.\cite{Hollingsworth2008-iy}  (We converted these
durations to rates like $1 / 2.90\;\text{month$^{-1}$} \approx
4.14\;\text{y$^{-1}$}$ and set the parameter using random samples from
$\Triangular(2, 4.14, 9.6)\;\text{y$^{-1}$}$ to capture the
uncertainty in its value.  See \autoref{uncertainty} for the
definition of $\Triangular(a, b, c)$ and other distributions.)  Acute
infection progresses into the chronic undiagnosed class ($U$), from
which diagnosis occurs at the rate $\alpha$, which is determined by
the country's current diagnosis level and its diagnosis target under
the different scenarios (\autoref{targets}).  Upon diagnosis, people
transition into the diagnosed class ($D$).  From the diagnosed class,
people transition to the treated compartment ($T$) upon starting ART
at the rate $\phi$, dependent on the country's current treatment level
and its treatment target (\autoref{targets}).  Treated people
transition to the viral suppression ($V$) at the rate $\gamma$,
sampled from $\Uniform(0.5, 1.5)\;\text{y$^{-1}$}$: thus, viral
suppression occurs after between 8 months and 2 years on
treatment.\cite{Currie2009-yz}  Disengagement from treatment moves
people from treated ($T$) and viral suppression ($V$) back to the
diagnosed class ($D$) at the rate $\psi$, dependent on the current
level of viral suppression and the country's target
(\autoref{targets}).

PLHIV without viral suppression (compartments $U$, $D$, and $T$)
develop AIDS (compartment $W$) after an average duration of 10.4
years,\cite{Morgan2002-cq} i.e.~with transition rate
$\sigma = 0.1064\;\text{y$^{-1}$}$.  Viral suppression greatly reduces
the rate of progression to AIDS ($\theta$) and improves longevity,
such that life expectancy is only $5$--$8$\;y shorter than average
country-specific life expectancy.\cite{Samji2013-kf, Unaids2014-ue}
This reduction in life expectancy was quantified by randomly sampling
the parameter $\omega$ from $\Uniform(5, 8)\;\text{y}$. Thus, in the
absence of relapse to untreated, the duration of viral suppression is
\begin{equation}
  \frac{1}{\theta + \mu} = \frac{1}{\mu} - \omega - \frac{1}{\nu}.
\end{equation}
The left-hand side of the equation is the time until leaving viral
suppression by either progression to AIDS (at rate $\theta$) or
background mortality (at rate $\mu$).  The right-hand side of the
equation is the life expectancy of uninfected people, minus the
reduction in life expectancy attributable to virally suppressed HIV
infection, minus the duration of the AIDS stage.  Solving for $\theta$
gives the rate of progression to AIDS from viral suppression,
\begin{equation}
  \label{theta}
  \theta = \frac{1}{1/\mu - \omega - 1/\nu} - \mu.
\end{equation}

@@

The background mortality rate ($\mu$) in each of the 127 countries was
collected from published
estimates.\cite{World_Development_Indicators2013-ee} The recruitment
rate ($\kappa$) into the age $15\;\text{y}$ and older group was set by
the difference between the country's population growth
rate\cite{WorldBankpg} and its background mortality
rate,\cite{World_Development_Indicators2013-ee} as would be true if
the population was at its stable age structure.  (More on these data
sources is in \autoref{data_sources}.)

@@

The relative transmission rates (per unit time), $\beta$, from acute
($A$), unsuppressed ($U$ and $T$), and suppressed infected ($V$)
people are respectively
\begin{equation}
  \label{betas}
  \begin{split}
    \beta_A &= 1 - (1 - \tau_A)^n,
    \\
    \beta_U &= 1 - (1 - \tau_U)^n,
    \\
    \beta_V &= 1 - (1 - \epsilon \tau_U)^n,
  \end{split}
\end{equation}
where the $\tau$'s are the relative transmissibilities per coital act,
$n$ is the annual number of coital acts, and $\epsilon$ is the
relative transmissibility per coital act with viral suppression.  For
transmissibility per coital act during the acute phase, we sampled
$\tau_A$ from
$\Triangular(0.0039, 0.0082,
0.0150)$.\cite{Wawer2005-us,
  Skarbinski2015-ni}
Transmissibility per coital act for unsuppressed people, $\tau_U$, was
sampled from
$\Triangular(0.00077, 0.0014, 0.00251)$.\cite{Hughes2012-so} The
transmissibility per coital act of people with viral suppression
relative to those without suppression has been estimated at 0.08 (95\%
confidence interval 0.002, 0.57).\cite{Donnell2010-xo} Due to the
extremely long left tail of this confidence interval, we sampled the
relative transmissibility per coital act $\epsilon$ from
$\BetaPERT(0.08, 0.002, 0.57)$ rather than a $\Triangular$
distribution.  (See \autoref{uncertainty} for a full description of
$\BetaPERT(a, b, c)$.)  The annual number of coital acts per person
$n$ was sampled from $\Uniform(96, 108)$ based on survey
studies.\cite{Wawer2005-us, Abdool_Karim2010-cm}  The country-specific
transmission rate $\xi$ was derived from country-specific longitudinal
trends in HIV prevalence and incidence, primarily from UNAIDS
estimates (\autoref{model_fitting}).

@@

The parametrized differential equations were solved numerically from
2015 to 2035 using the LSODA
routine\cite{odepack, scipy, medlock2016-git} and four effectiveness
outcomes were computed from the solutions.
\begin{description}[labelsep=0.6ex]
\item[Cumulative new infections] from 2015 to time $t$ was given by
  $Y(t)$.

@@

The results from the country simulations were aggregated to the
regional and global levels by adding the numbers of people in each
compartment over time and then new infections, per-capita incidence,
PLHIV, and HIV-related deaths were calculated from the aggregates.
(See Data Table 6 for the definitions of UNAIDS regions.)

@@

The proportion of people on ART who have achieved viral suppression is
\begin{equation}
  p_V(t) = \frac{V(t)}{T(t) + V(t)}.
\end{equation}
The rate of people relapsing to untreated is
\begin{equation}
  \label{relapse_rate}
  \psi(t) = \psi_{\max} H\big(p_V(t) - p_V^*(t)\big),
\end{equation}
with $\psi_{\max} = 1$, which, unlike diagnosis and treatment rates,
prevents relapses ($\psi(t) = 0$) when the current level of viral
suppression is below the target ($p_V^*(t) > p_V(t)$) and allows
relapses ($\psi(t) > 0$) when the current level of viral suppression
is at or above target.  We varied the relapse rate to achieve the
target level of viral suppression because achieving viral suppression
depends on the time on treatment\cite{Currie2009-yz}
(i.e.~$1 / \gamma$).

@@

We modeled 3 different targets for the proportion diagnosed ($p_D$), the
proportion treated ($p_T$), and the proportion with viral suppression ($p_V$).
\begin{description}[labelsep=0.6ex]
\item[Status quo] is maintenance of the proportions $p_D$, $p_T$, and
  $p_V$ at their initial levels going forward.  That is,
  \begin{equation}
    \label{status_quo_target}
    \begin{split}
      p_D^*(t) &= p_D(2015), \\
      p_T^*(t) &= p_T(2015), \\
      p_V^*(t) &= p_V(2015).
    \end{split}
  \end{equation}

@@

We found sufficient published estimates and data to parametrize our
model for 127 countries.  Demographic data, including population
growth rate,\cite{WorldBankpg} death
rate,\cite{World_Development_Indicators2013-ee} and number of people
aged 15--49 years\cite{The_World_Bank2016-fd} were obtained from the
World Bank (Data Table 2). Longitudinal HIV prevalence
(Data Table 3) and incidence (Data Table 4)
estimates for ages 15--49, spanning from as early as 1990 for some
countries to 2015, were primarily obtained from the AIDSinfo database
produced by UNAIDS.\cite{Unaids2016-an} Other sources included UNAIDS
Country Progress Reports\cite{Unaids2016-am} published from 2012 to
2016 and AIDS Data Hub,\cite{AIDSdatahub-fg} which compiles data from
UNAIDS, UNICEF, WHO, and the Asian Development Bank.  These sources
were also used to inform the initial conditions for the model
regarding the number of people diagnosed with HIV, the number on ART,
and the number who have viral suppression or have been retained on
treatment for at least 12 months (Data Table 5).  Where data
or estimates were not available from these sources, alternative
resources such as peer-reviewed journal articles and country health
ministry reports were consulted.  Full information on the sources used
for each country is provided in Data Tables 2--5.  These
source tables are also available in our public source-code
repository.\cite{medlock2016-git}

@@

We tailored our model to each of the 127 countries separately.  Using
available historical country-specific estimates of prevalence and
incidence (Data Tables 3 \& 4), and published estimates of
transmissibility per coital
act,\cite{Wawer2005-us, Donnell2010-xo, Hughes2012-so,
  Skarbinski2015-ni}
we derived an average country-specific transmission rate, taking into
account the differential contributions from the proportions of acute,
unsuppressed, and suppressed of the infected population in each
country.  Using historical data to smooth out short-term variations,
our calibration method provided estimates of current country-specific
transmission, as detailed below.

@@

Latin hypercube sampling\cite{blower1994} was used to draw parameter
samples from published parameter distributions (\autoref{model_param})
and the country-specific estimated distribution of the transmission
rate (\autoref{model_fitting}) 1000 times for each country.
Projections were then generated by running the model simulations with
these parameter sets.  To compare interventions, as well as status quo
projections, we implemented the variance-reduction technique of using
the same set of sample parameter values for all interventions,
repeating this process for each of the 1000 simulations in every
country.\cite{shechter2006}  We report the median, 1st and 3rd
quartiles (i.e.~25th and 75th percentiles), and 5th and 95th
percentiles for each model outcome.  We also calculated partial rank
correlation coefficients (PRCC)\cite{blower1994} to measure the
independent effect of each non-vaccine parameter on model projections
of global HIV incidence from 2015 to 2035 (Extended Data
Fig.~\ref*{PRCCs}).

>>>>

\usepackage{amsmath}
\usepackage{graphics}
\usepackage{tikz}

% Tweak spacing of descriptions.
\usepackage{enumitem}
\setlist[description]{topsep=1ex, labelsep=0.6ex}

% Tweak captions.
\usepackage[font=sf,
            labelfont=bf,
            labelsep=period,
            singlelinecheck=off]{caption}

% Make float text and math sans serif.
\usepackage[font=sf]{floatrow}
\usepackage{sansmath}
\usepackage{etoolbox}
\AtBeginEnvironment{table}{\sansmath}
\AtBeginEnvironment{figure}{\sansmath}
\floatsetup[table]{capposition=top,  % Table captions on top.
                   justification=justified}  % Don't center by default.

% Nice formatting for \textwidth tables.
\usepackage{tabularx}

\usepackage[noblocks]{authblk}
\renewcommand{\Affilfont}{\small}

@@

% \usepackage[style=nature]{biblatex}
% \addbibresource{supporting_information.bib}
\usepackage[numbers, round, comma, sort&compress]{natbib}
\usepackage{hypernat}
% PNAS style: no brackets in reference list.
\makeatletter 
\renewcommand\@biblabel[1]{#1.} 
\makeatother

% PNAS style: square brackets around equation labels and all in
% boldface, boldface equation number in references.
\usepackage{mathtools}
\newtagform{brackets}{\bf[}{]}
\usetagform{brackets}
\renewcommand{\eqref}[1]{\textbf{\ref{#1}}}

\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\newcommand{\sectionname}{Section}
\renewcommand{\sectionautorefname}{\sectionname}
\renewcommand{\figurename}{Fig.}
\renewcommand{\figureautorefname}{\figurename}

\usepackage{xr}
\externaldocument{extended_data_fig_1}

\newcommand{\md}{\mathrm{d}}
\DeclareMathOperator{\Uniform}{Uniform}
\DeclareMathOperator{\Triangular}{Triangular}
\DeclareMathOperator{\Lognormal}{Lognormal}
\DeclareMathOperator{\BetaPERT}{Beta-PERT}

@@

\author[a,1]{Jan Medlock}
\affil[a]{Department of Biomedical Sciences,
  Oregon State University,
  106 Dryden Hall,
  Corvallis, OR 97331-4801,
  USA}
\author[b]{Abhishek Pandey}
\author[b]{Alyssa S.~Parpia}
\author[b]{Amber Tang}
\author[b]{Laura A.~Skrip}
\author[b]{Alison P.~Galvani}
\affil[b]{Center for Infectious Disease Modeling and Analysis,
  Yale School of Public Health,
  135 College Street,
  New Haven, CT 06510-2483,
  USA}
\affil[1]{To whom correspondence should be addressed.
  Email: \href{mailto:jan.medlock@oregonstate.edu}
  {\texttt{jan.medlock@oregonstate.edu}}.}

@@

We developed a mathematical model of HIV transmission and progression
that stratifies HIV infection into acute, chronic undiagnosed, chronic
diagnosed, chronic treated, chronic virally suppressed, and AIDS,
along with uninfected unvaccinated and vaccinated people
(\autoref{model} \& \autoref{model_diag}), tailored to
each of 127 countries.  Simulations of our model project the number of
people in each HIV stratum from 2015 to 2035, from which statistics
such as prevalence and incidence were calculated.  To evaluate the
impact of the 90--90--90 and 95--95--95 UNAIDS goals, model transition
rates were adjusted dynamically to meet intervention goals for
diagnosis, treatment, viral suppression, and vaccination coverage
(\autoref{targets}).  Data and estimates from many sources, including
The World Bank and UNAIDS, were used to parametrize the model
(\autoref{data_sources}).  Country-specific transmission rates were
fitted to estimates of the historical trajectories of incidence and
prevalence, spanning from as early as 1990 for some countries
(\autoref{model_fitting}). For every scenario of intervention
combinations, we conducted 1000 model simulations, sampling values of
model parameters from empirical distributions for each simulation, and
reported the results with median and percentiles
(\autoref{uncertainty}).

\begin{figure}
  \centering
  \input{diagram}
  \caption{Model diagram.}
  \label{model_diag}
\end{figure}

The model simulation and analysis tools, written in Python, are
publicly available \cite{medlock2016-git}.

@@

Within each country, our model divides people aged $15$\;years and
older into non-overlapping HIV states.  We focused on this age group
because it accounts for the vast majority of global PLHIV
\cite{UNICEF}, 93\% in 2014; because population-level HIV estimates
and demographic data are typically only estimated for ages $15$--$49$;
and to capture the continuing burden of infections in people ages $50$
and older.  Our model has 8 HIV states (\autoref{model_diag}):
susceptible to HIV infection ($S$ is the number of people
susceptible), vaccinated against HIV ($R$), acute HIV infection ($A$),
undiagnosed HIV infection ($U$), diagnosed but untreated HIV infection
($D$), treated without viral suppression ($T$), viral suppression
($V$), and having AIDS ($W$).

@@

\begin{table}
  \caption{Model parameters.}
  \label{model_param}
  \begin{tabularx}{\textwidth}{lXlll}
    \hline
    & Definition & Value & References
    \\ \hline
    $\mu$ & Background death rate
    & Country specific, Data Set 2
    & \cite{World_Development_Indicators2013-ee}
    \\
    $\kappa$ & Recruitment rate
    & Country specific, Data Set 2
    & \cite{World_Development_Indicators2013-ee, WorldBankpg}
    \\
    $\delta$ & Rate of progression from acute infection
    & $\Triangular(2, 4.14, 9.6)\;\text{y$^{-1}$}$
    & \cite{Hollingsworth2008-iy}
    \\
    $\sigma$ & Rate of developing AIDS without viral suppression
    & $0.1064\;\text{y$^{-1}$}$ & \cite{Morgan2002-cq}
    \\
    $\theta$ & Rate of developing AIDS with viral suppression
    & Eq.~\eqref{theta} & ---
    \\
    $\gamma$ & Rate of viral suppression
    & $\Uniform(0.5, 1.5)\;\text{y$^{-1}$}$
    & \cite{Currie2009-yz}
    \\
    $\nu$ & Death rate from AIDS & $0.5\;\text{y$^{-1}$}$
    & \cite{Morgan2002-cq}
    \\
    $\rho$ & Vaccine efficacy & 50\%\,[30\%, 70\%] & ---
    \\
    $\omega$ & Reduction in life expectancy with viral suppression
    & $\Uniform(5, 8)\;\text{y}$
    & \cite{Samji2013-kf, Unaids2014-ue}
    \\
    $\tau_{A}$ & Transmissibility per coital act during acute phase
    & $\Triangular(0.0039, 0.0082, 0.0150)$
    & \cite{Wawer2005-us, Skarbinski2015-ni}
    \\
    $\tau_{U}$ & Transmissibility per coital act after acute phase
    & $\Triangular(0.00077, 0.0014, 0.00251)$
    & \cite{Hughes2012-so}
    \\
    $\epsilon$ & Relative transmissibility per coital act with
    viral suppression & $\BetaPERT(0.08, 0.002, 0.57)$
    & \cite{Donnell2010-xo, attia_2009, wilson_2012, jia_2013,
      rodger_2016}
    \\
    $n$ & Coital acts per year & $\Uniform(96, 108)$
    & \cite{Wawer2005-us, Abdool_Karim2010-cm}
    \\
    $\alpha$ & Diagnosis rate & Eq.~\eqref{diagnosis_rate} & ---
    \\
    $\phi$ & Treatment rate & Eq.~\eqref{treatment_rate} & ---
    \\
    $\psi$ & Rate of relapse to untreated & Eq.~\eqref{relapse_rate} & ---
    \\
    $\zeta$ & Vaccination rate & Eq.~\eqref{vaccination_rate} & ---
    \\ \hline
  \end{tabularx}
  \\[5pt]
  $\Triangular$, $\Uniform$, and $\BetaPERT$ are sampling
  distributions (\autoref{uncertainty}).
  \\
  For vaccine efficacy, $50\%$ was the baseline, and $30\%$ and $70\%$
  were also assessed in sensitivity analysis.
\end{table}

People susceptible to HIV ($S$) are at risk of becoming infected
according to the force of infection ($\lambda$), which depends on the
HIV transmission rate, estimated from the country-specific incidence
and prevalence data (\autoref{model_fitting}), adjusted for the
current number of acutely infected, untreated or ineffectively
treated, and virally suppressed PLHIV.  Upon infection, people
transition into acute infection ($A$), characterized by high viral
loads for an average of 2.90 [95\% confidence interval 1.23,
6.00]\;months \cite{Hollingsworth2008-iy}.  (We converted these
durations to rates like $1 / 2.90\;\text{month$^{-1}$} \approx
4.14\;\text{y$^{-1}$}$ and set the parameter using random samples from
$\Triangular(2, 4.14, 9.6)\;\text{y$^{-1}$}$ to capture the
uncertainty in its value.  See \autoref{uncertainty} for the
definition of $\Triangular(a, b, c)$ and other distributions.)  Acute
infection progresses into the chronic undiagnosed class ($U$), from
which diagnosis occurs at the rate $\alpha$, which is determined by
the country's current diagnosis level and its diagnosis target under
the different scenarios (\autoref{targets}).  Upon diagnosis, people
transition into the diagnosed class ($D$).  From the diagnosed class,
people transition to the treated compartment ($T$) upon starting ART
at the rate $\phi$, dependent on the country's current treatment level
and its treatment target (\autoref{targets}).  Treated people
transition to the viral suppression ($V$) at the rate $\gamma$,
sampled from $\Uniform(0.5, 1.5)\;\text{y$^{-1}$}$: thus, viral
suppression occurs after between 8 months and 2 years on
treatment \cite{Currie2009-yz}.  Disengagement from treatment moves
people from treated ($T$) and viral suppression ($V$) back to the
diagnosed class ($D$) at the rate $\psi$, dependent on the current
level of viral suppression and the country's target
(\autoref{targets}).

PLHIV without viral suppression (compartments $U$, $D$, and $T$)
develop AIDS (compartment $W$) after an average duration of 10.4
years \cite{Morgan2002-cq}, i.e.~with transition rate
$\sigma = 0.1064\;\text{y$^{-1}$}$.  Viral suppression greatly reduces
the rate of progression to AIDS ($\theta$) and improves longevity,
such that life expectancy is only $5$--$8$\;y shorter than average
country-specific life expectancy \cite{Samji2013-kf, Unaids2014-ue}.
This reduction in life expectancy was quantified by randomly sampling
the parameter $\omega$ from $\Uniform(5, 8)\;\text{y}$. Thus, in the
absence of relapse to untreated, the duration of viral suppression is
\begin{equation}
  \frac{1}{\theta + \mu} = \frac{1}{\mu} - \omega - \frac{1}{\nu}.
\end{equation}
The left-hand side of the equation is the time until leaving viral
suppression by either progression to AIDS (at rate $\theta$) or
background mortality (at rate $\mu$).  The right-hand side of the
equation is the life expectancy of uninfected people, minus the
reduction in life expectancy attributable to virally suppressed HIV
infection, minus the duration of the AIDS stage.  Solving for $\theta$
gives the rate of progression to AIDS from viral suppression,
\begin{equation}
  \label{theta}
  \theta = \frac{1}{1/\mu - \omega - 1/\nu} - \mu.
\end{equation}

@@

The background mortality rate ($\mu$) in each of the 127 countries was
collected from published
estimates \cite{World_Development_Indicators2013-ee}.  The recruitment
rate ($\kappa$) into the age $15\;\text{y}$ and older group was set by
the difference between the country's population growth
rate \cite{WorldBankpg} and its background mortality
rate \cite{World_Development_Indicators2013-ee}, as would be true if
the population was at its stable age structure.  (More on these data
sources is in \autoref{data_sources}.)

@@

The relative transmission rates (per unit time), $\beta$, from acute
($A$), unsuppressed ($U$ and $T$), and suppressed infected ($V$)
people are respectively
\begin{equation}
  \label{betas}
  \begin{split}
    \beta_A &= 1 - (1 - \tau_A)^n,
    \\
    \beta_U &= 1 - (1 - \tau_U)^n,
    \\
    \beta_V &= 1 - (1 - \epsilon \tau_U)^n,
  \end{split}
\end{equation}
where the $\tau$'s are the relative transmissibilities per coital act,
$n$ is the annual number of coital acts, and $\epsilon$ is the
relative transmissibility per coital act with viral suppression.  For
transmissibility per coital act during the acute phase, we sampled
$\tau_A$ from
$\Triangular(0.0039, 0.0082,
0.0150)$ \cite{Wawer2005-us,
  Skarbinski2015-ni}.
Transmissibility per coital act for unsuppressed people, $\tau_U$, was
sampled from
$\Triangular(0.00077, 0.0014, 0.00251)$ \cite{Hughes2012-so}.  The
transmissibility per coital act of people with viral suppression
relative to those without suppression has been estimated at 0.08 (95\%
confidence interval 0.002, 0.57) \cite{Donnell2010-xo}.  Due to the
extremely long left tail of this confidence interval, we sampled the
relative transmissibility per coital act $\epsilon$ from
$\BetaPERT(0.08, 0.002, 0.57)$ rather than a $\Triangular$
distribution.  (See \autoref{uncertainty} for a full description of
$\BetaPERT(a, b, c)$.)  The annual number of coital acts per person
$n$ was sampled from $\Uniform(96, 108)$ based on survey
studies \cite{Wawer2005-us, Abdool_Karim2010-cm}.  The country-specific
transmission rate $\xi$ was derived from country-specific longitudinal
trends in HIV prevalence and incidence, primarily from UNAIDS
estimates (\autoref{model_fitting}).

@@

The parametrized differential equations were solved numerically from
2015 to 2035 using the LSODA
routine \cite{odepack, scipy, medlock2016-git} and four effectiveness
outcomes were computed from the solutions.
\begin{description}
\item[Cumulative new infections] from 2015 to time $t$ was given by
  $Y(t)$.

@@

The results from the country simulations were aggregated to the
regional and global levels by adding the numbers of people in each
compartment over time and then new infections, per-capita incidence,
PLHIV, and HIV-related deaths were calculated from the aggregates.
(See Data Set 6 for the definitions of UNAIDS regions.)

@@

The proportion of people on ART who have achieved viral suppression is
\begin{equation}
  p_V(t) = \frac{V(t)}{T(t) + V(t)}.
\end{equation}
The rate of people relapsing to untreated is
\begin{equation}
  \label{relapse_rate}
  \psi(t) = \psi_{\max} H\big(p_V(t) - p_V^*(t)\big),
\end{equation}
with $\psi_{\max} = 1$, which, unlike diagnosis and treatment rates,
prevents relapses ($\psi(t) = 0$) when the current level of viral
suppression is below the target ($p_V^*(t) > p_V(t)$) and allows
relapses ($\psi(t) > 0$) when the current level of viral suppression
is at or above target.  We varied the relapse rate to achieve the
target level of viral suppression because achieving viral suppression
depends on the time on treatment \cite{Currie2009-yz}.

@@

We modeled 3 different targets for the proportion diagnosed ($p_D$), the
proportion treated ($p_T$), and the proportion with viral suppression ($p_V$).
\begin{description}
\item[Status quo] is maintenance of the proportions $p_D$, $p_T$, and
  $p_V$ at their initial levels going forward.  That is,
  \begin{equation}
    \label{status_quo_target}
    \begin{split}
      p_D^*(t) &= p_D(2015), \\
      p_T^*(t) &= p_T(2015), \\
      p_V^*(t) &= p_V(2015).
    \end{split}
  \end{equation}

@@

We found sufficient published estimates and data to parametrize our
model for 127 countries.  Demographic data, including population
growth rate \cite{WorldBankpg}, death
rate \cite{World_Development_Indicators2013-ee}, and number of people
aged 15--49 years \cite{The_World_Bank2016-fd} were obtained from the
World Bank (Data Set 2). Longitudinal HIV prevalence
(Data Set 3) and incidence (Data Set 4)
estimates for ages 15--49, spanning from as early as 1990 for some
countries to 2015, were primarily obtained from the AIDSinfo database
produced by UNAIDS \cite{Unaids2016-an}. Other sources included UNAIDS
Country Progress Reports \cite{Unaids2016-am} published from 2012 to
2016 and AIDS Data Hub \cite{AIDSdatahub-fg}, which compiles data from
UNAIDS, UNICEF, WHO, and the Asian Development Bank.  These sources
were also used to inform the initial conditions for the model
regarding the number of people diagnosed with HIV, the number on ART,
and the number who have viral suppression or have been retained on
treatment for at least 12 months (Data Set 5).  Where data
or estimates were not available from these sources, alternative
resources such as peer-reviewed journal articles and country health
ministry reports were consulted.  Full information on the sources used
for each country is provided in Data Sets 2--5.  These
source tables are also available in our public source-code
repository \cite{medlock2016-git}.

@@

We tailored our model to each of the 127 countries separately.  Using
available historical country-specific estimates of prevalence and
incidence (Data Sets 3 \& 4), and published estimates of
transmissibility per coital
act \cite{Wawer2005-us, Donnell2010-xo, Hughes2012-so,
  Skarbinski2015-ni},
we derived an average country-specific transmission rate, taking into
account the differential contributions from the proportions of acute,
unsuppressed, and suppressed of the infected population in each
country.  Using historical data to smooth out short-term variations,
our calibration method provided estimates of current country-specific
transmission, as detailed below.

@@

Latin hypercube sampling \cite{blower1994} was used to draw parameter
samples from published parameter distributions (\autoref{model_param})
and the country-specific estimated distribution of the transmission
rate (\autoref{model_fitting}) 1000 times for each country.
Projections were then generated by running the model simulations with
these parameter sets.  To compare interventions, as well as status quo
projections, we implemented the variance-reduction technique of using
the same set of sample parameter values for all interventions,
repeating this process for each of the 1000 simulations in every
country \cite{shechter2006}.  We report the median, 1st and 3rd
quartiles (i.e.~25th and 75th percentiles), and 5th and 95th
percentiles for each model outcome.  We also calculated partial rank
correlation coefficients (PRCC) \cite{blower1994} to measure the
independent effect of each non-vaccine parameter on model projections
of global HIV incidence from 2015 to 2035 (\autoref{PRCCs}).

\begin{figure}
   \centering
   \includegraphics{../Codes/plots/prcc.pdf}
   \caption{Sensitivity of global HIV incidence from 2015 to 2035 to
     model parameters.  Left is the sensitivity of infections averted
     by 95--95--95, relative to status quo.  Right is the sensitivity
     of vaccination with status quo levels of diagnosis and treatment,
     relative to status quo.  Sensitivity is measured by the partial
     rank correlation coefficient (PRCC).}
   \label{PRCCs}
\end{figure}
